---
title: "Customer_segmenation in r"
output: html_document
---

```{r}
## importing data reading it
customer_data=read.csv("Mall_Customers.csv")

str(customer_data)

names(customer_data)

head(customer_data)

summary(customer_data)
```

```{r}

#sd of all names
sd(customer_data$Age)
summary(customer_data$Annual.Income..k..)
sd(customer_data$Annual.Income..k..)
summary(customer_data$Age)
```

```{r}
##bar plot visualization for column names

a=table(customer_data$Gender)
barplot(a,main="Using BarPlot to display Gender Comparision",
        ylab="Count",
        xlab="Gender",
        col=rainbow(2),
        legend=rownames(a))
```

##conclusion = no of females are higher than males

```{r}

##pie chart to observer the ratio of male and female

pct=round(a/sum(a)*100)
lbs=paste(c("Female","Male")," ",pct,"%",sep=" ")
install.packages("plotrix")
library(plotrix)
pie3D(a,labels=lbs,main="Pie Chart Depicting Ratio of Female and Male")

```

#conlcusion that percentage of female is 56% whereas male is as 44 %

```{r}
#now visualization of age of the people
summary(customer_data$Age)
#plotting a histogram
hist(customer_data$Age,
     col="blue",
     main="Histogram to Show Count of Age Class",
     xlab="Age Class",
     ylab="Frequency",
     labels=TRUE)
```

```{r}
#boxplot 
boxplot(customer_data$Age,
        col="blue",
        main="Boxplot for Descriptive Analysis of Age")

```

#col=max age is b/w 30 and 35 the min age is 18 and max is 70

```{r}

#summary(customer_data$Annual.Income..k..)
hist(customer_data$Annual.Income..k..,
     col="#660033",
     main="Histogram for Annual Income",
     xlab="Annual Income Class",
     ylab="Frequency",
     labels=TRUE)
```

```{r}
#density plot
plot(density(customer_data$Annual.Income..k..),
     col="yellow",
     main="Density Plot for Annual Income",
     xlab="Annual Income Class",
     ylab="Density")
polygon(density(customer_data$Annual.Income..k..),
        col="#ccff66")
```

## con=min income is 15 and max is 137 averaage income of 70

```{r}

##spending score of the customes
summary(customer_data$Spending.Score..1.100.)


boxplot(customer_data$Spending.Score..1.100.,
        horizontal=TRUE,
        col="red",
        main="BoxPlot for Descriptive Analysis of Spending Score")
#histogram 
hist(customer_data$Spending.Score..1.100.,
     main="HistoGram for Spending Score",
     xlab="Spending Score Class",
     ylab="Frequency",
     col="#6600cc",
     labels=TRUE)

```

##applying k mean clustring algorithm

#for k using the elbo method

```{r}
#install.packages("tidyverse")
library(purrr)
set.seed(123)

# function to calculate total intra-cluster sum of square 
customer_data
customer_data[,3:5]

iss <- function(k) {
  kmeans(customer_data[,3:5],k,iter.max=100,nstart=100,algorithm="Lloyd" )$tot.withinss
}

k.values <- 1:10


iss_values <- map_dbl(k.values, iss)

plot(k.values, iss_values,
     type="b", pch = 19, frame = FALSE, 
     xlab="Number of clusters K",
     ylab="Total intra-clusters sum of squares")
```

##Average Silhouette Method

```{r}
library(cluster) 
#install.packages("gridExtra")
library(gridExtra)
library(grid)

# when k=2
k2<-kmeans(customer_data[,3:5],2,iter.max=100,nstart=50,algorithm="Lloyd")
s2<-plot(silhouette(k2$cluster,dist(customer_data[,3:5],"euclidean")))


# when k=3
k3<-kmeans(customer_data[,3:5],3,iter.max=100,nstart=50,algorithm="Lloyd")
s3<-plot(silhouette(k3$cluster,dist(customer_data[,3:5],"euclidean")))
# when k=4
k4<-kmeans(customer_data[,3:5],4,iter.max=100,nstart=50,algorithm="Lloyd")
s4<-plot(silhouette(k4$cluster,dist(customer_data[,3:5],"euclidean")))
#when k=5
k5<-kmeans(customer_data[,3:5],5,iter.max=100,nstart=50,algorithm="Lloyd")
s5<-plot(silhouette(k5$cluster,dist(customer_data[,3:5],"euclidean")))
# when k=6
k6<-kmeans(customer_data[,3:5],6,iter.max=100,nstart=50,algorithm="Lloyd")
s6<-plot(silhouette(k6$cluster,dist(customer_data[,3:5],"euclidean")))
# 7when k=
k7<-kmeans(customer_data[,3:5],7,iter.max=100,nstart=50,algorithm="Lloyd")
s7<-plot(silhouette(k7$cluster,dist(customer_data[,3:5],"euclidean")))
# when k=
k8<-kmeans(customer_data[,3:5],8,iter.max=100,nstart=50,algorithm="Lloyd")
s8<-plot(silhouette(k8$cluster,dist(customer_data[,3:5],"euclidean")))


# now vizuaising the cluster 
#install.packages("NbClust")
library(NbClust)
#install.packages("colorspace")
#install.packages("factoextra")
library(factoextra)

fviz_nbclust(customer_data[,3:5], kmeans, method = "silhouette")

```

##Gap Statistic Method

```{r}
set.seed(125)
stat_gap <- clusGap(customer_data[,3:5], FUN = kmeans, nstart = 25,
                    K.max = 10, B = 50)
fviz_gap_stat(stat_gap)

```

######now taking 6 as optimal clusters

```{r}
k6<-kmeans(customer_data[,3:5],6,iter.max=100,nstart=50,algorithm="Lloyd")
k6


#Visualizing the Clustering Results using the First Two Principle Components
pcclust=prcomp(customer_data[,3:5],scale=FALSE) #principal component analysis
summary(pcclust)

pcclust$rotation[,1:2]
```

## annual income and spending score

```{r}
set.seed(1)
ggplot(customer_data, aes(x =Annual.Income..k.., y = Spending.Score..1.100.)) + 
  geom_point(stat = "identity", aes(color = as.factor(k6$cluster))) +
  scale_color_discrete(name=" ",
                       breaks=c("1", "2", "3", "4", "5","6"),
                       labels=c("Cluster 1", "Cluster 2", "Cluster 3", "Cluster 4", "Cluster 5","Cluster 6")) +
  ggtitle("Segments of Mall Customers", subtitle = "Using K-means Clustering")

```

#spending score and age

```{r}


ggplot(customer_data, aes(x =Spending.Score..1.100., y =Age)) + 
  geom_point(stat = "identity", aes(color = as.factor(k6$cluster))) +
  scale_color_discrete(name=" ",
                       breaks=c("1", "2", "3", "4", "5","6"),
                       labels=c("Cluster 1", "Cluster 2", "Cluster 3", "Cluster 4", "Cluster 5","Cluster 6")) +
  ggtitle("Segments of Mall Customers", subtitle = "Using K-means Clustering")

```

#classes vs clusters

```{r}
kCols=function(vec){cols=rainbow (length (unique (vec)))
return (cols[as.numeric(as.factor(vec))])}

digCluster<-k6$cluster; dignm<-as.character(digCluster); # K-means clusters

plot(pcclust$x[,1:2], col =kCols(digCluster),pch =19,xlab ="K-means",ylab="classes")
legend("bottomleft",unique(dignm),fill=unique(kCols(digCluster)))



```
